A Hitchhiker's Guide to Reproducible Research in R

### 2021-07-14 manuscript.Rmd
better guide readers through the advanced section closes #145 closes #147

Some projects, however, are not as easy to make reproducible.
As explained above, repro is merely a simplified interface that guarantuees a high level of reproducibility for simple projects.
These simplifications are possible because of two restrictions.
First, the project can only rely on R packages for the Dockerfile and second, it can have no multilayered preprocessing for the Makefile.<!--AB: please explain both points in more detail-->
Lifting these restrictions requires the user to interact more directly with Make or Docker.
To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.

@@

With this approach, users of `repro` can not only build complex software environments but they can also implement complex file dependencies.
The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
If you can formulate the creation of dependencies in terms of computational steps you should include these in the Makefile.<!--AB I do not understand this sentence-->
The Makefile that `repro` creates is only an template and you are free to change it.
However, make sure to never remove the following two lines:

>>>>

Some projects, however, are not as easy to make reproducible.
As explained above, `repro` is merely a simplified interface that guarantees a high level of reproducibility for simple projects.
These simplifications are possible because of two restrictions.
Users who ask themselves either "How can I install software that is not an R package in the Docker image?" or "How can I express complex dependencies between files, e.g. raw data turned into preprocessed data required by analysis?" need to be aware of these restriction and require a deeper understanding of the inner workings of `repro`.
All other users may safely skip this section and may return to it if they become limited by restriction.

The metadata the `repro::automate()` function relies on can only express dependencies R packages for the Dockerfile and only trivial dependencies (in the form of "file must exist") for the Makefile. 
The first limitation means that users must rely on software that is either already provided by the base Dockerimage "rocker/verse" or the R packages they list in the metadata.
Other software that users might need like other programming languages, not already installed latex packages etc. have to be added manually.
We plan to add support for commonly used ways to install software beyond R packages via the metadata and `repro::automate()` e.g. for system libraries (via `apt` the Ubuntu package manager), Latex packages (via `tlmgr` the Tex Live package manager), python packages (via `pip` the python package manager).
The second limitation is about dependencies.
Make can represent complex dependencies like in this example: A depends on B, which in turn depends on C and D.
If B is missing in this example, Make would know how to recreate it from C and D.
These dependencies and especially how these dependencies should be resolved are difficult to represent in the metadata.
Users therefore have to either "flatten" the dependency structure, by simply stating that A depends on B, C and D therefore leaving out important information or express the dependencies directly within the Makefile.

The following section explains how to overcome these limitation while still relying on the automatons afforded by `repro`.
Lifting these restrictions requires the user to interact more directly with Make or Docker.
To satisfy more complicated requirements users need to understand how repro utilizes Make and Docker internally.

@@

With this approach, users of `repro` can not only build complex software environments but they can also implement complex file dependencies.
The standard `repro` metadata only makes sure that all dependencies are available but does not allow you to specify custom recipes for them in the metadata.
If you can formulate the creation of dependencies in terms of computational steps, e.g. the file `data/clean.csv` is created from `data/raw.csv` by script `R/preprocess.R`, you should include these in the Makefile.
The Makefile that `repro` creates is only an template and you are free to change it.
However, make sure to never remove the following two lines:
